---
title: "Creating Genome Alignments for TOGA"
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: false
    toc_float: true
    toc_collapsed: false
    theme: default
    css: "theme3.css"
toc_depth: 1
number_sections: true
editor_options:
  chunk_output_type: console
---

## make_lastz_chains {.tabset}

### What is make_lastz_chains?

>This program is also from Hiller Lab (see [TOGA](toga.html)) and it is a pipeline that takes in a reference and a query genome either as .fna or .2bit files and creates pairwise genome alignment pairs from them for use in TOGA. 

Read more in depth on their GitHub [here](https://github.com/hillerlab/make_lastz_chains).

### Installation and Running

This program is supported on MacOS, but the Hiller Lab team *highly recommends* running it on a Linux-based machine.

#### On Grace
This pipeline is already installed in Grace, but I am not sure how to use it yet. Will update (eventually).

#### In a Conda Environment
Follow the instructions on the Hiller Lab GitHub [here](https://github.com/hillerlab/make_lastz_chains) to install.

### Troubleshooting
####  fatal: Could not read from remote repository. Please make sure you have the correct access rights and the repository exists.

The full error is here:
```{r remote-fatal-fail, eval=FALSE}
Warning: Permanently added the ECDSA host key for IP address '140.82.113.3' to the list of known hosts. Permission denied (publickey). fatal: Could not read from remote repository. Please make sure you have the correct access rights and the repository exists.
```

I corrected this by going into my home > .ssh directory in Grace and adding the passkey from *id_rsa.pub* to my GitHub SSH keys. This is the public key the Grace terminal uses to download files from GitHub. This allowed me to download from the Hiller Lab repository.  

If your *id_rsa.pub* file is empty, visit [this link](https://www.theserverside.com/blog/Coffee-Talk-Java-News-Stories-and-Opinions/Fix-GitHubs-Permission-denied-publickey-SSH-error). It has the full documentation that I followed to fix this error.

#### Error: detected space-or-tab-containing sequence

The whole error is below:

```{r space-error, eval=FALSE}
Error! File: data/Nitens-and-Greg/GCF_021461395.2_iqSchAmer2.1_genomic.fna - detected space-or-tab-containing sequence:
NC_060119.1 Schistocerca americana isolate TAMUIC-IGC-003095 chromosome 1, iqSchAmer2.1, whole genome shotgun sequence
Please exclude or fix sequences with spaces and tabs.
```

I have submitted this error to the Hiller Lab GitHub and will update when I get a response.
