---
title: "Mapping Transcription Factors to Trancription Factor Families"
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: false
    toc_float: true
    toc_collapsed: false
    theme: default
    css: "theme3.css"
toc_depth: 1
number_sections: true
editor_options:
  chunk_output_type: console
---

## Mapping Transcription Factors {.tabset}


### Running on a Cluster
<center>
<mark> This process is configured to run on the ASU Sols cluster. </mark>
</center>

To conduct this process we will be using HMMER to map our transcription factors against the Pfam database. Additionally, this tutorial will be done the proteins of *Schistocerca piceifrons*. The protein data is available from NCBI [here](https://ftp.ncbi.nlm.nih.gov/genomes/all/GCF/021/461/385/GCF_021461385.2_iqSchPice1.1/)

#### Step 1: Download and Prepare Pfam

```{r, eval=FALSE}

# i) Get the latest Pfam-A HMMs
wget ftp://ftp.ebi.ac.uk/pub/databases/Pfam/current_release/Pfam-A.hmm.gz
gunzip Pfam-A.hmm.gz

module load hmmer-3.1b2-gcc-12.1.0 

# ii) Index for hmmscan
hmmpress Pfam-A.hmm

# iii) Download protein data from NCBI, decompress file, and rename
wget https://ftp.ncbi.nlm.nih.gov/genomes/all/GCF/021/461/385/GCF_021461385.2_iqSchPice1.1/GCF_021461385.2_iqSchPice1.1_protein.faa.gz

gunzip GCF_021461385.2_iqSchPice1.1_protein.faa.gz

mv GCF_021461385.2_iqSchPice1.1_protein.faa GCF_021461385.2_iqSchPice1.1_protein.fasta

```

#### Step 2: Run hmmscan on proteins

```{r, eval=FALSE}

hmmscan --cpu 4 \
  --domtblout pfam.domtblout \
  Pfam-A.hmm \
  GCF_021461385.2_iqSchPice1.1_protein.fasta

```

This step will create a file called *pfam.domtblout*. Use the command ls -lh to check the file is not empty and the step has successfully run.

#### Step 3: Parse and map to transcription factor families in R

Create a file named parse-map-tf.R and copy and paste the following code into it:

```{r, eval = FALSE}
library(dplyr)

# 1) Read the domtblout (skip "#" comments)
dom <- read.delim("pfam.domtblout",
                  header=FALSE,
                  sep="")

# 2) Name the first few columns per HMMER spec
colnames(dom)[1:19] <- c(
  "GeneID","Accession","Domain","EValue","Score","Bias",
  paste0("i",1:10),
  "EnvStart","EnvEnd","DomainLength"
)

# 3) Define which Pfam accessions → TF Family
#    (extend this list with all families you care about)
pfam2fam <- tibble(
  Accession = c("PF22861.1","PF00046","PF02365","PF00249","PF04434"),
  Family    = c("GEO12453p1-like",       "GATA",    "NAC",    "MYB",    "TCP")
)

# 4) Filter only hits to those TF domains
tf_hits <- dom %>%
  filter(Accession %in% pfam2fam$Accession) %>%
  distinct(GeneID, Accession)

# 5) Join to get Family names
reg_fam <- tf_hits %>%
  left_join(pfam2fam, by = "Accession") %>%
  select(GeneID, Family) %>%
  distinct()

# 6) Your regulator vector
regulators <- reg_fam$GeneID

# 7) Quick sanity check
cat("Found", nrow(reg_fam), "putative TFs across", length(unique(reg_fam$Family)), "families\n")
table(reg_fam$Family)

```

Then create a file called run-parse-map.sh and copy and paste the following code into it:

```{r, eval=FALSE}
#!/bin/bash

##NECESSARY JOB SPECIFICATIONS
#SBATCH --job-name=parse-and-map-piceifrons       #Set the job name to "psmc-referencegenome"
#SBATCH --time=02:00:00         #Set the wall clock limit to 2 hours
#SBATCH --ntasks=1             #Request 1 task
#SBATCH --cpus-per-task=8       #Request 8 cpus
#SBATCH --mem=30G              #Request 30GB per node
#SBATCH --output=stdout.%x.%j
#SBATCH --error=stderr.%x.%j

module load r-4.3.2-gcc-11.2.0

Rscript parse-map-tf.R

```

After this has completed running, use the command *cat stdout.{YOUR_JOB_NUMBER}* to check the results. If you followed this tutorial, the output should read:

```{r, eval=FALSE}

Found 1 putative TFs across 1 families

GEO12453p1-like 
              1 
```

Don't worry if you know there's multiple proteins in this family, it is only checking for unique protein families not individual proteins.

### Troubleshooting

#### Found 0 putative TFs across 0 families
For me, this error occured when no proteins in my pfam2fam matched proteins in the pfam.domtblout file. To verify this is the issue, use the command *head pfam.domtblout* and choose the first protein listed. Rerun the program and it should say "Found 1 putative TFs across 1 families" along with the protein family listed below this message.
