---
title: "De Novo Transcriptome Assembly"
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: false
    toc_float: true
    toc_collapsed: false
    theme: default
    css: "theme3.css"
toc_depth: 1
number_sections: true
editor_options:
  chunk_output_type: console
---

## De Novo Transcriptome Assembly {.tabset}


### Data Preprocessing

#### i) Load Necessary Software
##### ia) Load R Modules

Before we begin processing our files, we need to load the following R libraries in our R session. Load the R module using the command "module load r-4.3.2-gcc-11.2.0" and copy and paste the following code into your R session:

```{r, eval=FALSE}

library("knitr")
library("rmdformats")
library("tidyverse")
library("DT")  # for making interactive search table
library("plotly") # for interactive plots
library("ggthemes") # for theme_calc
library("reshape2")
library("readr")
library("ggplot2")

## Global options
options(max.print="10000")
knitr::opts_chunk$set(
    echo = TRUE,
    message = FALSE,
    warning = FALSE,
    cache = FALSE,
    comment = FALSE,
    prompt = FALSE,
    tidy = TRUE
)
opts_knit$set(width=75)

```


##### ib) Create Mamba Environment for MultiQC

The package MultiQC is not readily available on the Sol cluster, so we will need to install it via a Mamba environment. Log into the Sol cluster via ssh and follow the below commands to create a new Mamba environment and install the MultiQC package into this environment.

```{r, eval = FALSE}

interactive -p htc -c 4 -t 30

module load mamba/latest

mamba create -n trans-map -c conda-forge -c bioconda python=3 multiqc trimgalore

```

After it finishes loading, you will be prompted to answer yes or no (Y/n), type Y to accept.

To then verify the creation of this environment, type the command:

```{r, eval=FALSE}

source activate trans-map

python

import multiqc

```

If you receive no errors, then installation is successful!

#### ii) Quality Control

First, we will need to assess each .fastq file's quality using FASTQC. Create a file called *qual-control.sh* and copy and paste the follwoing code into it:

```{r, eval=FALSE}
#!/bin/bash

##NECESSARY JOB SPECIFICATIONS
#SBATCH --job-name=quality-control       #Set the job name to "psmc-referencegenome"
#SBATCH --time=02:00:00         #Set the wall clock limit to 8 hours
#SBATCH --ntasks=1             #Request 1 task
#SBATCH --cpus-per-task=8       #Request 8 cpus
#SBATCH --mem=30G              #Request 30GB per node
#SBATCH --mail-user={USER@EMAIL.COM}
#SBATCH --output=stdout.%x.%j
#SBATCH --error=stderr.%x.%j

module load fastqc-0.11.3-gcc-12.1.0

mkdir fastqc_out

fastqc -t 8 ../data/raw/*.fastq -o ../data/fastqc_out/

module purge

module load mamba/latest 

source activate trans-map

multiqc --title 'MultiQC Report' -v /scratch/ebaker48/greg_transcript/data/fastqc_out

```

Edit the email in the slurm header to your email (if you wish to receive email alerts when your job begins and ends, otherwise delete that line), the path to your fastqc_out directory (your current directory + /fastqc_out/), and optionally, the title of the MultiQC report.

This produces FASTQC report files for each fastq file and the MULTIQC is a compiles report summarizing all of the FASTQC reports.

#### iii) Trim and Remove Adapters

To perform this trimming step we will use trim_galore to remove the contaminants from the files.

Create a file called *trim.sh* and copy and paste the following code into it:

```{r, eval=FALSE}

#!/bin/bash
#SBATCH --job-name=trim_galore_all
#SBATCH --output=trim_galore_all_%j.log
#SBATCH --error=trim_galore_all_%j.err
#SBATCH --time=12:00:00
#SBATCH --mem=32G
#SBATCH -c 8

# Load Trim Galore via mamba environment (Step 0)
module purge
module load mamba/latest
source activate trans-map

INDIR="../data/raw"
OUTDIR="../data/trimmed"

# --------------------------------------------------
# Loop through paired-end FASTQ files
# --------------------------------------------------
for READ1 in "${INDIR}"/*_R1_001.fastq; do
    READ2="${READ1/_R1_001.fastq/_R2_001.fastq}"

    if [[ -f "$READ2" ]]; then
        echo "Processing: $READ1 and $READ2"

        trim_galore --trim-n \
            --cores 8 \
            --quality 20 \
            --clip_R1 2 \
            --clip_R2 2 \
            --nextera \
            --fastqc \
            --output_dir "$OUTDIR" \
            --paired "$READ1" "$READ2"

    else
        echo "Warning: Could not find pair for $READ1"
    fi
done


```

#### iv) Trimming Quality Check

To ensure our trimming was not overly aggressive, we can perform one last quality check of our trimmed files using the below script:

```{r, eval=FALSE}

#!/bin/bash
#SBATCH --job-name=fastqc_trimmed
#SBATCH --output=fastqc_trimmed_%j.log
#SBATCH --error=fastqc_trimmed_%j.err
#SBATCH --time=3:00:00
#SBATCH --mem=500M
#SBATCH --cpus-per-task=2

module load fastqc-0.11.3-gcc-12.1.0

fastqc ../data/trimmed/*_val_1.fq -o ../data/trimmed/
fastqc ../data/trimmed/*_val_2.fq -o ../data/trimmed/

```








### De Novo Assembly 
